MAESTROeX: A Massively Parallel Low Mach Number Astrophysical Solver

### 2019-03-13 MAESTROeX_I/paper.tex
first cut at etarho and split projection discussion

In this step, using equation (\ref{eq:momentum}), 
we compute time-centered edge velocities, $\uadvonedag$, using
$\Ub^n$.  The $\dagger$ superscript refers to the
fact that the predicted velocity field does not satisfy the divergence
constraint.  We then construct $\uadvone$ from $\uadvonedag$
using a MAC projection.
We note that $\uadvone$ satisfies the divergence constraint
\begin{equation}
\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left(S^{\nph,\star} - \frac{1}{\gammaonebar^np_0^n}\frac{\partial p_0}{\partial t} \right).
\end{equation}
In practice we solve for.

@@

The procedure to construct $\uadvtwodag$ is identical to the Godunov procedure
for computing $\uadvonedag$ in {\bf Step 3}, but uses
the updated value $S^{\nph}$ rather than $S^{\nph,\star}$.
We note that $\uadvtwo$ satisfies the divergence constraint
\begin{equation}
\nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
\beta_0^{\nph} \left(S^{\nph} - \frac{1}{\gammaonebar^{n+\myhalf}p_0^{n+\myhalf}}\frac{\partial p_0}{\partial t}\right),
\end{equation}
where
\begin{equation}
\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2},
\qquad
\gammaonebar^{\nph} = \frac{ \gammaonebar^n +  \gammaonebar^{n+1,\star} }{2}.
\qquad
p_0^{\nph} = \frac{ p_0^n + p_0^{n+1,\star} }{2},
\end{equation}

@@

Finally, we use an approximate nodal projection to define $\Ub^{n+1}$
from $\Ub^{n+1,\dagger},$  such that $\Ub^{n+1}$ approximately
satisfies
\begin{equation}
\nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right) = \beta_0^{n+1} \left( S^{n+1} - \frac{1}{\gammaonebar^{n+1}p_0^{n+1}}\frac{\partial p_0}{\partial t} \right).
\end{equation}
As part of the projection we also define the new-time perturbational pressure,
$\pi^\nph.$  This projection necessarily differs from the MAC projection used in
{\bf Step 3} and {\bf Step 7} because the velocities in those steps are defined
on edges and $\Ub^{n+1}$ is defined at cell centers, requiring different divergence
and gradient operators.  Details of the approximate projection are given in Paper III.

@@

\appendix
\section{Projection Details}

>>>>

In this step, using equation (\ref{eq:momentum}), 
we compute time-centered edge velocities, $\uadvonedag$, using
$\Ub^n$.  The $\dagger$ superscript refers to the
fact that the predicted velocity field does not satisfy the divergence
constraint.  We then construct $\uadvone$ from $\uadvonedag$
using a MAC projection.
We note that $\uadvone$ satisfies the divergence constraint
\begin{equation}
\nabla \cdot \left(\beta_0^n \uadvone\right) = \beta_0^n \left[S^{\nph,\star} - \frac{1}{\gammaonebar^np_0^n}\left(\frac{\partial p_0}{\partial t}\right)^{n-\myhalf} \right].\label{eq:div1}
\end{equation}
In practice we solve for.

@@

The procedure to construct $\uadvtwodag$ is identical to the Godunov procedure
for computing $\uadvonedag$ in {\bf Step 3}, but uses
the updated value $S^{\nph}$ rather than $S^{\nph,\star}$.
We note that $\uadvtwo$ satisfies the divergence constraint
\begin{equation}
\nabla \cdot \left(\beta_0^{\nph} \uadvtwo\right) =
\beta_0^{\nph} \left[S^{\nph} - \frac{1}{\gammaonebar^{n+\myhalf}p_0^{n+\myhalf}}\left(\frac{\partial p_0}{\partial t}\right)^{n+\myhalf}\right],\label{eq:div2}
\end{equation}
where
\begin{equation}
\beta_0^{\nph} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2},
\qquad
\gammaonebar^{\nph} = \frac{ \gammaonebar^n +  \gammaonebar^{n+1,\star} }{2}.
\qquad
p_0^{\nph} = \frac{ p_0^n + p_0^{n+1,\star} }{2},
\end{equation}

@@

Finally, we use an approximate nodal projection to define $\Ub^{n+1}$
from $\Ub^{n+1,\dagger},$  such that $\Ub^{n+1}$ approximately
satisfies the divergence constraint,
\begin{equation}
\nabla \cdot \left(\beta_0^{n+1} \Ub^{n+1} \right) = \beta_0^{n+1} \left[ S^{n+1} - \frac{1}{\gammaonebar^{n+1}p_0^{n+1}}\left(\frac{\partial p_0}{\partial t}\right)^{n+\myhalf}\right].\label{eq:div3}
\end{equation}
As part of the projection we also define the new-time perturbational pressure,
$\pi^\nph.$  This projection necessarily differs from the MAC projection used in
{\bf Step 3} and {\bf Step 7} because the velocities in those steps are defined
on edges and $\Ub^{n+1}$ is defined at cell centers, requiring different divergence
and gradient operators.  Details of the approximate projection are given in Paper III.

@@

\appendix
\section{Projection Details}
To enforce the divergence constraint in {\bf Step 3, 7}, and {\bf 11}, we use a projection method analgous to the methods originally developed for incompressible flow.

Based on our past experience in the MAESTRO project, we have found it useful to split the velocity dynamics into a perturbational and base state velocity,
\begin{equation}
\Ub = \Ubt(\xb,t) + w_0(r,t)\eb_r,
\end{equation}
solve for each term separately, and immediately combine them to find a full velocity that satisfies the constraint.  We take that approach here, primarily because it allows us to enforce a boundary condition on $w_0$ at the edge of the star (i.e., the cutoff density location where we hold density constant).  Namely, to enforce that $r^2 w_0$ remain constant is difficult to do when solving for the full velocity.  So in practice, we solve the constraint over the lateral average,
\begin{equation}
\nabla\cdot\left(\beta_0\Ubt\right) = \beta_0(S-\overline{S})
\end{equation}
and sepratately solve for $w_0$ using 
\begin{equation}
\nabla\cdot(\beta_0w_0\eb_r) = \beta_0\left(\overline{S} - \frac{1}{\gammaonebar p_0}\frac{\partial p_0}{\partial t}\right).
\end{equation}
To find $w_0$ we integrate in 1D using the procedure in Appendix B of \cite{MAESTRO_V}.

Note that in this approach, we estimated the time-derivative of the pressure in part by examining how laterally averaged $\rho'=\rho-\rho_0$ changed over time (quantified by $\eta_\rho = \overline{\rho'\Ub\cdot\eb_r}$.  After evolving the species, we compute this term in the same way as reported in \cite{MAESTRO_V} for use in computing $w_0$.
